### 2022-05-20 examples/benchmark/full-report.qmd
update

---
title: Inventor Disambiguation Report
format:
    html:
        theme: cosmo
        code-fold: true
        code-tools: true
        code-block-border-left: true
        code-copy: true
        code-block-border-left: "#31BAE9"
jupyter: python3
---

## Data Preparation

First, we prepare the disambiguation result data to evaluate. We consider the current PatentsView disambiguation, obtained from the `rawinventor.tsv` data file, as well as one trivial disambiguation where all inventor mentions are clustered together.

The code below reads the disambiguation result from `rawinventor.tsv`. A disambiguation result is defined as a pandas Series with an index representing mention IDs and values representing cluster assignement.

Mention IDs are standardized to be of the form `<OFFICE><PATENT_NUMBER>-<INVENTOR_SEQUENCE>`, where OFFICE is the patent issuer office abbreviation (e.g. "US" or "EPO"), PATENT_NUMBER is the standard patent number form, and INVENTOR_SEQUENCE is the inventor's position on the given patent, starting from 0. For example, the mention ID
> US5811073-0
represents the first listed inventor on US patent number 5811073.

```{python}
#| label: data-prep
#| code-fold: true
import pandas as pd

import wget
import zipfile
import os

import plotly.io as pio
pio.templates.default = "plotly_white" # Set plotly theme

if not os.path.isfile("rawinventor.tsv"):
    wget.download("https://s3.amazonaws.com/data.patentsview.org/download/rawinventor.tsv.zip")
    with zipfile.ZipFile("rawinventor.tsv.zip", 'r') as zip_ref:
        zip_ref.extractall(".")
    os.remove("rawinventor.tsv.zip")

if not os.path.isfile("disambiguation.tsv"):
    rawinventor = pd.read_csv("rawinventor.tsv", sep="\t")
    rawinventor["mention-id"] = "US" + rawinventor.patent_id.astype(str) + "-" + rawinventor.sequence.astype(str)
    rawinventor[["mention-id", "inventor_id"]].to_csv("disambiguation.tsv", sep="\t", index=False)
    del(rawinventor)

data = pd.read_csv("disambiguation.tsv", sep="\t")
data.set_index("mention-id", inplace=True)
current_disambiguation = data["inventor_id"]
```

Next, we define a dictionary describing the two disambiguations to compare, i.e. the "current" disambiguation and the "one cluster" trivial disambiguation.

```{python}
#| label: disambiguations
#| code-fold: false

disambiguations = {
    "current": current_disambiguation,
    "one cluster": 0*current_disambiguation
}
```

We can now generate the disambiguation report using `disambiguations` and `rawinventor.tsv` as the input.

## Summary Statistics

```{python}
#| label: imports

import wget
import zipfile
import os

import plotly.io as pio
pio.templates.default = "plotly_white" # Set plotly theme

from pv_evaluation.summary import InventorDisambiguationSummary

summarizer = InventorDisambiguationSummary("rawinventor.tsv", processed_data_dir=".", name="current disambiguation")
```

```{python}
summarizer.plot_cluster_size_distribution(range=(0.5,10.5))
```

```{python}
summarizer.get_top_inventors()
```

```{python}
summarizer.plot_entropy_curve()
```

```{python}
summarizer.plot_cluster_unique_name_distribution(range=(0, 100))
```

```{python}
summarizer.plot_homonymy_rate_distribution()
```

## Benchmark Evaluation