### 2022-05-23 examples/benchmark/full-report.qmd
update

---
title: Inventor Disambiguation Report
date: today
author: PatentsView-Evaluation
toc: true
toc-location: left
format:
    html:
        theme: cosmo
        echo: false
        fig-cap-location: margin
        code-fold: true
        code-tools: true
        code-line-numbers: true
        code-block-border-left: true
        code-copy: true
        code-block-border-left: "#31BAE9"
        code-link: true
        include-in-header: header.html
        include-after-body: footer.html
execute:
  cache: true
  freeze: true
  warning: false
  message: false
jupyter: python3
---

<!--
## Data Preparation

First, we prepare the disambiguation result data to evaluate. We consider the current PatentsView disambiguation, obtained from the `rawinventor.tsv` data file, as well as one trivial disambiguation where all inventor mentions are clustered together.

The code below reads the disambiguation result from `rawinventor.tsv`. A disambiguation result is defined as a pandas Series with an index representing mention IDs and values representing cluster assignement.

Mention IDs are standardized to be of the form `<OFFICE><PATENT_NUMBER>-<INVENTOR_SEQUENCE>`, where OFFICE is the patent issuer office abbreviation (e.g. "US" or "EPO"), PATENT_NUMBER is the standard patent number form, and INVENTOR_SEQUENCE is the inventor's position on the given patent, starting from 0. For example, the mention ID
> US5811073-0
represents the first listed inventor on US patent number 5811073.
-->
```{python}
#| label: data-prep
#| code-fold: true
import pandas as pd

import wget
import zipfile
import os

import plotly.io as pio
pio.templates.default = "plotly_white" # Set plotly theme

if not os.path.isfile("rawinventor.tsv"):
    wget.download("https://s3.amazonaws.com/data.patentsview.org/download/rawinventor.tsv.zip")
    with zipfile.ZipFile("rawinventor.tsv.zip", 'r') as zip_ref:
        zip_ref.extractall(".")
    os.remove("rawinventor.tsv.zip")

if not os.path.isfile("disambiguation.tsv"):
    rawinventor = pd.read_csv("rawinventor.tsv", sep="\t")
    rawinventor["mention-id"] = "US" + rawinventor.patent_id.astype(str) + "-" + rawinventor.sequence.astype(str)
    rawinventor[["mention-id", "inventor_id"]].to_csv("disambiguation.tsv", sep="\t", index=False)
    del(rawinventor)

data = pd.read_csv("disambiguation.tsv", sep="\t")
data.set_index("mention-id", inplace=True)
current_disambiguation = data["inventor_id"]
```

<!--
Next, we define a dictionary describing the two disambiguations to compare, i.e. the "current" disambiguation and the "one cluster" trivial disambiguation.-->

```{python}
#| label: disambiguations
#| code-fold: false

disambiguations = {
    "current": current_disambiguation,
    "one cluster": 0*current_disambiguation
}
```

<!--
We can now generate the disambiguation report using `disambiguations` and `rawinventor.tsv` as the input.
-->

## Summary Statistics

```{python}
#| label: imports

import wget
import zipfile
import os

import plotly.io as pio
pio.templates.default = "plotly_white" # Set plotly theme

from pv_evaluation.summary import InventorDisambiguationSummary
from pv_evaluation.benchmark import inventor_benchmark_plot

summarizer = InventorDisambiguationSummary("rawinventor.tsv", processed_data_dir=".", name="current disambiguation")
```

::: {.panel-tabset}

### Cluster Sizes

```{python}
#| label: cluster-size-distribution-plot
#| fig-cap: This is a test caption
#| cap-location: margin
summarizer.plot_cluster_size_distribution(range=(0.5,10.5))
```

:::{.callout-tip collapse="true"}

## Detailed Explanation (click to expand)

The above plot provides information about the number of inventors with a given number of (co-)authored patents.

We can read from the plot the number of inventors with a single (co-)authored patent, with exactly two (co-)authored patents, and so forth. This distribution of the number of (co-)authored patents patents per inventor is equivalently called the **cluster sizes distribution** of the disambiguation.

When comparing disambiguation results, look for major shifts in the cluster size distribution. Is one of the distribution more skewed to the left than another? This could indicate that one of the disambiguation favors smaller clusters, possibly resulting in higher precision but lower recall.

:::

### Top Inventors

```{python}
#summarizer.get_top_inventors()
```

### Entropy Curve

```{python}
#summarizer.plot_entropy_curve()
```

### Cluster Homogeneity

```{python}
#summarizer.plot_cluster_unique_name_distribution(range=(0, 100))
```

### Cluster Separation

```{python}
#summarizer.plot_homonymy_rate_distribution()
```

:::

## Benchmark Evaluation

::: {.panel-tabset}

### Evaluation Results

```{python}
inventor_benchmark_plot(disambiguations)
```

### Predicted Cluster Errors

::: {.panel-tabset}

#### Israeli Inventors

```{python}
#| column: page

from IPython.core.display import display, HTML

from pv_evaluation.benchmark import (
    load_israeli_inventors_benchmark,
    load_patentsview_inventors_benchmark,
    inspect_clusters_to_split,
    inspect_clusters_to_merge,
    style_cluster_inspection,
)

def style_and_display(table, by="prediction"):
    return display(HTML(style_cluster_inspection(table.reset_index(), by=by).hide_index().to_html()))

style_and_display(inspect_clusters_to_split(disambiguations["current"], load_israeli_inventors_benchmark()), by="prediction")
```

#### PatentsView Inventors

```{python}
#| column: page
style_and_display(inspect_clusters_to_split(disambiguations["current"], load_patentsview_inventors_benchmark()), by="prediction")
```

:::

### Reference Cluster Errors

::: {.panel-tabset}

#### Israeli Inventors

```{python}
#| column: page
style_and_display(inspect_clusters_to_merge(disambiguations["current"], load_israeli_inventors_benchmark()), by="reference")
```

#### PatentsView Inventors

```{python}
#| column: page
style_and_display(inspect_clusters_to_merge(disambiguations["current"], load_patentsview_inventors_benchmark()), by="reference")
```

:::

:::

## Representative Performance Estimates