### 2022-05-23 pv_evaluation/templates/inventor/callouts/cluster-homogeneity-tips.qmd

:::{.callout-tip collapse="true"}

## Detailed Explanation (click to expand)

Cluster homogeneity is the level similarity of among a cluster's elements. For inventor disambiguation, this is about the variation between the way that an inventor is represented on its patents (e.g., different name spellings).

Here we look at cluster homogeneity from a binary perspective -- whether or not there is variation, within a cluster, in how an inventor's name is spelled. The proportion of inventors with a unique name (no name variation within its cluster) is our metric of cluster homogeneity called the **homogeneity rate**.

In the plot below, the homogeneity rate (i.e., clusters with no name variation) is plotted as a function of cluster size. For inventors with a single patent, the proportion of homogeneous clusters is trivially 1. For inventors with two patents, we can read off the proportion of them with no name variation, and so forth.

When comparing two disambiguation results, look for changes in the homogeneity rate across cluster sizes. A higher homogeneity rate means possibly smaller, more robust clusters. On the other hand, lower homogeneity rates may be associated with an increased error probability.

:::