### 2022-05-25 pv_evaluation/templates/inventor/callouts/cluster-errors-tips.qmd
add callout tips

:::{.callout-tip collapse="true"}

## Detailed Explanation (click to expand)

Cluster error tabs (predicted cluster errors and reference cluster errors) allow you to investigate clusters which contain errors.

**Predicted cluster errors** are about predicted clusters which should have been split up in two or more parts, according to the reference data.

To investigate predicted cluster errors, sort the table by `prediction` id and look for mismatching `reference` ids within predicted clusters.

**Reference cluster errors are about predicted clusters which should be merged together, according to the reference data.

To investigate reference cluster errors, sort the table by `reference` id and look for mismatching `cluster` ids within reference clusters.

:::