### 2022-12-02 pv_evaluation/templates/inventor/0-report.qmd
update

{% include 'inventor/1-header.qmd' %}

This is report has been automatically generated for the following data source(s) at {{ datetime }}:

{% for filename in disambiguation_files %}
- {{ filename }}
{% endfor %}

For more information, please refer to the [PatentsView-Evaluation project homepage](https://github.com/PatentsView/PatentsView-Evaluation/) or to the report [source code](index.qmd).

{% include 'inventor/2-summary.qmd' %}

{% include 'inventor/3-benchmark.qmd' %}